\documentclass[11pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\signature{Heitor Oliveira Gonçalves, Ph.D. Candidate\\
\small Graduate Program in Computational Modeling\\
\small Catholic University of Petrópolis (UCP)\\
\small Email: heitorhog@gmail.com}

\address{Graduate Program in Computational Modeling\\
Catholic University of Petrópolis (UCP)\\
Rua Benjamin Constant 213, Centro\\
Petrópolis, RJ, 25610-130, Brazil}

\begin{document}

\begin{letter}{Editor-in-Chief\\
Journal of Computing in Civil Engineering\\
American Society of Civil Engineers (ASCE)}

\opening{Dear Editor,}

We are pleased to submit our manuscript entitled \textbf{``Automated Corrosion Detection in ASTM A572 Grade 50 W-Beams Using U-Net and Attention U-Net: A Comparative Analysis for Semantic Segmentation''} for consideration for publication in the \textit{Journal of Computing in Civil Engineering}.

\textbf{Summary:} This study presents a rigorous comparative analysis between U-Net and Attention U-Net architectures for automated corrosion detection in ASTM A572 Grade 50 steel W-beams. Using a dataset of 414 annotated images, we demonstrate that Attention U-Net achieves statistically superior performance with mean IoU of 0.775 compared to 0.693 for classical U-Net (11.8\% improvement, p < 0.001, Cohen's d = 0.98).

\textbf{Key Contributions:}
\begin{itemize}
    \item First rigorous comparative evaluation between U-Net and Attention U-Net specifically for corrosion detection in ASTM A572 Grade 50 structural steel
    \item Comprehensive statistical analysis with significance testing and effect size quantification
    \item Attention map interpretability analysis providing insights into model decision-making
    \item Practical implementation guidelines for field deployment
\end{itemize}

\textbf{Relevance to JCCE:} This research directly addresses the journal's scope by advancing computational methods for structural inspection. The methodology bridges artificial intelligence and civil engineering, providing practical tools for infrastructure maintenance programs. The 46\% reduction in false positives and improved detection of incipient corrosion have significant economic implications for structural health monitoring.

\textbf{Declarations:}
\begin{itemize}
    \item This manuscript has not been published previously and is not under consideration for publication elsewhere.
    \item All authors have approved the manuscript and agree with its submission.
    \item There are no conflicts of interest to declare.
    \item The research did not involve human subjects or animals.
\end{itemize}

We believe our work represents a significant contribution to the field and aligns well with the readership of the \textit{Journal of Computing in Civil Engineering}. We look forward to your favorable consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
